---
title: "cbm_exn_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libcbm cbm_exn package example in R

The following example shows how to run `libcbm` `cbm_exn` in R.  `cbm_exn` is a package that supports running the CBM-CFS3 Carbon dynamics driven by net above ground forest Carbon increments of Merchantable, Foliage and Other pools.  This is in contrast to the default CBM-CFS3 method of generating increments using the CBM3 volume to biomass routine.

```{r}
library(reticulate)
library(plyr)
#json <- import("json")
cbm_exn_model <- import("libcbm.model.cbm_exn.cbm_exn_model")
libcbm_resources <- import("libcbm.resources")
model_variables <- import("libcbm.model.model_definition.model_variables")
output_processor <- import("libcbm.model.model_definition.output_processor")
```

## spinup and step functions

```{r}
spinup <- function(spinup_input, parameters) {
  with(cbm_exn_model$initialize(parameters=parameters) %as% cbm, {
      # the spinup function creates the t=0 cbm_vars
      # but you can save and or load cbm_vars for each
      # timestep at the end of spinup point
      cbm_vars = cbm$spinup(spinup_input)
      return (cbm_vars)
    }
  )
}

step <- function(cbm_vars, parameters) {

    with(cbm_exn_model$initialize(parameters=parameters) %as% cbm, {
      cbm_vars = cbm$step(cbm_vars)
      return (cbm_vars)
    }
  )
}
```

## load net increments

Load some net increments based on a CBM-CFS3 tutorial1 run

```{r}
net_increments <- read.csv(
  file.path(
    libcbm_resources$get_test_resources_dir(),
    "cbm_exn_net_increments",
    "net_increments.csv"
  )
)
```

repeat the net increments to illustrate simulation of multiple stands

```{r}
colnames(net_increments) <- c("age", "merch_inc", "foliage_inc", "other_inc")
stand_increments <- NULL
n_stands <- 1000
for(i in 0:(n_stands-1)){
  copied_increments <- data.frame(net_increments)
  copied_increments <- cbind(data.frame(row_idx=i), copied_increments)
  stand_increments = rbind(
    stand_increments, copied_increments
  )
}

```

set up some spin-up parameters

```{r}
spinup_parameters <- data.frame(
  age=sample(0L:60L, n_stands, replace=T),
  area=rep(1.0, n_stands),
  delay=rep(0L, n_stands),
  return_interval=rep(125L, n_stands),
  min_rotations=rep(10L, n_stands),
  max_rotations=rep(30L, n_stands),
  spatial_unit_id=rep(17L, n_stands), # Ontario/Mixedwood plains
  species=rep(20L, n_stands), # red pine
  mean_annual_temperature=rep(2.55, n_stands),
  historical_disturbance_type=rep(1L, n_stands),
  last_pass_disturbance_type=rep(1L, n_stands)
)

```

## assemble parameters
```{r}
# this is the path to some default bundled parameters for cbm_exn
# any method for building equivalently formatted dataframes will also work
param_path <- libcbm_resources$get_cbm_exn_parameters_dir()
cbm_exn_parameters <- dict(
    # TODO: need a solution for loading json that works correctly
    #pools=json$loads((paste(readLines(file.path(param_path, "pools.json")), collapse = " "))),
    #flux=json$loads((paste(readLines(file.path(param_path, "flux.json")), collapse = " "))),
    slow_mixing_rate=read.csv(file.path(param_path, "slow_mixing_rate.csv")),
    turnover_parameters=read.csv(file.path(param_path, "turnover_parameters.csv")),
    species=read.csv(file.path(param_path, "species.csv")),
    root_parameters=read.csv(file.path(param_path, "root_parameters.csv")),
    decay_parameters=read.csv(file.path(param_path, "decay_parameters.csv")),
    disturbance_matrix_value=read.csv(file.path(param_path, "disturbance_matrix_value.csv")),
    disturbance_matrix_association=read.csv(file.path(param_path, "disturbance_matrix_association.csv"))
)
```

## run spinup
```{r}
cbm_vars = spinup(
  dict(
    parameters=spinup_parameters,
    increments=stand_increments
  ),
  cbm_exn_parameters
)
```


## run 50 timesteps

```{r}

# helper to store results in-memory
out_processor = output_processor$ModelOutputProcessor()

for(t in 1:50){
  cbm_vars$parameters$mean_annual_temperature = 2.55
  cbm_vars$parameters$disturbance_type = sample(
    c(0L, 1L, 4L), n_stands, replace=T, prob=c(0.98, 0.01, 0.01)
  )

  # look up the original increments and join to the current stand age
  step_increments = join(
    x=data.frame(age=cbm_vars$state$age),
    y=net_increments,
    by="age")

  # since some of the ages are out of range for the defined
  # data, set the increments to 0
  step_increments$merch_inc[is.na(step_increments$merch_inc)] = 0
  step_increments$foliage_inc[is.na(step_increments$foliage_inc)] = 0
  step_increments$other_inc[is.na(step_increments$other_inc)] = 0

  #assign the merged increments to the parameters dataframe
  cbm_vars$parameters$merch_inc = step_increments$merch_inc
  cbm_vars$parameters$foliage_inc = step_increments$foliage_inc
  cbm_vars$parameters$other_inc = step_increments$other_inc
  cbm_vars = step(cbm_vars, cbm_exn_parameters)

  out_processor$append_results(
    t, model_variables$ModelVariables$from_pandas(cbm_vars)
  )
}

```

```{r}
results <- out_processor$get_results()

```

Due to how `Reticulate` works, the following are actually R data.frames
```{r}

pools <- results["pools"]$to_pandas()
flux <- results["flux"]$to_pandas()
parameters <- results["parameters"]$to_pandas()
state <- results["state"]$to_pandas()
```

```{r}
pools
```

```{r}
state
```