---
title: "cbm3_tutorial2.rmd"
output: html_document
---

``` {r}
library(ggplot2)
library(reshape2)

library(reticulate)

# if necessary specify a particular python
#use_python("/usr/local/bin/python")

os <- import("os")
sit_cbm_factory <- import("libcbm.input.sit.sit_cbm_factory")
cbm_simulator <- import("libcbm.model.cbm.cbm_simulator")
cbm_output <- import("libcbm.model.cbm.cbm_output")
resources <- import("libcbm.resources")
```

## Setup

Load the standard import tool configuration at the specified path. This 
configuration encompasses the data source for the various sit inputs
(sit_inventory, sit_classifiers etc.) and also the relationships of classifiers,
and disturbance types to the default CBM parameters.

``` {r}
config_path <- os$path$join(
  resources$get_test_resources_dir(), "cbm3_tutorial2", "sit_config.json")
sit <- sit_cbm_factory$load_sit(config_path)
```

Initialize and validate the inventory in the sit dataset

``` {r}
inv_tuple <- sit_cbm_factory$initialize_inventory(sit)
classifiers <- inv_tuple[[1]]
inventory <- inv_tuple[[2]]
```


``` {r}
cbm_output_processor <- cbm_output$CBMOutput()
```

## Simulation


``` {r}
with(sit_cbm_factory$initialize_cbm(sit) %as% cbm, {

  rule_based_processor = sit_cbm_factory$create_sit_rule_based_processor(sit, cbm)
  
  cbm_simulator$simulate(
      cbm,
      n_steps              = 200,
      classifiers          = classifiers,
      inventory            = inventory,
      pre_dynamics_func    = rule_based_processor$pre_dynamics_func,
      reporting_func       = cbm_output_processor$append_simulation_result
  )
  
})
```

## Pool Results

``` {r}

biomass_pools = c('SoftwoodMerch','SoftwoodFoliage', 'SoftwoodOther', 'SoftwoodCoarseRoots', 'SoftwoodFineRoots',
                 'HardwoodMerch', 'HardwoodFoliage', 'HardwoodOther', 'HardwoodCoarseRoots', 'HardwoodFineRoots')

dom_pools = c('AboveGroundVeryFastSoil', 'BelowGroundVeryFastSoil', 'AboveGroundFastSoil', 'BelowGroundFastSoil',
             'MediumSoil', 'AboveGroundSlowSoil', 'BelowGroundSlowSoil', 'SoftwoodStemSnag', 'SoftwoodBranchSnag',
             'HardwoodStemSnag', 'HardwoodBranchSnag')

pools_df = cbm_output_processor$pools$to_pandas()
df <- data.frame(
  timestep = pools_df["timestep"],
  biomass = rowSums(pools_df[biomass_pools]),
  dom = rowSums(pools_df[dom_pools]),
  total_eco = rowSums(pools_df[c(biomass_pools, dom_pools)])
)

df <- aggregate(
  cbind(biomass=df$biomass, dom=df$dom, total_eco=df$total_eco),
  by=list(timestep=df$timestep), FUN=sum)

df <- melt(df, id.vars="timestep", variable.name="series")
ggplot(df, aes(timestep,value)) + geom_line(aes(colour = series))

```

