---
title: "multi stand level modelling in libcbm (R version)"
output: html_notebook
---

```{r}
library(reticulate)
```

Optional Set up steps: only need to run the following section when python is not
already set up

```{r}
version <- "3.11:latest"
install_python(version)
virtualenv_create("my-environment", version = version)
# install libcbm and its dependencies
py_install("libcbm", "my-environment")
```

Activate a python environment named "my-environment"

```{r}
use_virtualenv("my-environment")
```

python imports

```{r}
np = import("numpy")
libcbm_dataframe = import("libcbm.storage.dataframe")
stand_cbm_factory = import("libcbm.model.cbm.stand_cbm_factory")
cbm_output = import("libcbm.model.cbm.cbm_output")
cbm_simulator = import("libcbm.model.cbm.cbm_simulator")
cbm_variables = import("libcbm.model.cbm.cbm_variables")
cbm_defaults = import ("libcbm.model.cbm.cbm_defaults")
resources = import("libcbm.resources")
cbm_defaults_ref = import("libcbm.model.cbm.cbm_defaults_reference")
```

Define merchantable volume and classifiers input used to drive CBM.

The values shown here are purely for illustrative purposes.

Inventory and yields have a m:1 relationship, and the relationship is defined by
the classifier values attached to each merch volume curve and each inventory
record.

This information is well-suited for storage in a `json` formatted file in order
to scale up this information.

```{r}

classifiers = dict(
    c1=list("c1_v1", "c1_v2"),
    c2=list("c2_v1")
)


merch_volumes = list(
    dict(
        classifier_set=list("c1_v1", "?"),
        merch_volumes=list(
          dict(
            species="Spruce",
            age_volume_pairs=list(
                list(0, 0),
                list(50, 100),
                list(100, 150),
                list(150, 200)
            )
          ),
          dict(
            species="Oak",
            age_volume_pairs=list(
                list(0, 0),
                list(50, 100),
                list(100, 150),
                list(150, 200)
            )
          )
        )
    ),
    dict(
        classifier_set=list("c1_v2", "c2_v1"),
        merch_volumes=list(
          dict(
            species="Spruce",
            age_volume_pairs=list(
                list(0, 0),
                list(50, 100),
                list(100, 150),
                list(150, 200)
            )
          ),
          dict(
            species="Oak",
            age_volume_pairs=list(
                list(0, 0),
                list(50, 100),
                list(100, 150),
                list(150, 200)
            )
          )
        )
    )
)

```

Define example inventory, this could also be derived from CSV.

```{r}
n_steps=50
n_stands=1000

inventory = libcbm_dataframe$from_pandas(
  data.frame(
    c1="c1_v1",
    c2="c2_v1",
    admin_boundary="Ontario",
    eco_boundary="Mixedwood Plains",
    age=sample(0L:60L, n_stands, replace=T),
    area=1.0,
    delay=0L,
    land_class="UNFCCC_FL_R_FL",
    afforestation_pre_type="None",
    historic_disturbance_type="Wildfire",
    last_pass_disturbance_type="Wildfire"

  )
)
```


Default parameters and defaults reference: inspection and modification of CBM
parameters

```{r}
defaults_ref <- cbm_defaults_ref$CBMDefaultsReference(
  resources$get_cbm_defaults_path()
)
```

Example showing how to interact with one of the properties of `defaults_ref`

```{r}
defaults_ref$as_data_frame(defaults_ref$disturbance_type_ref)

```

```{r}
default_parameters <- cbm_defaults$get_cbm_parameters_factory(
  resources$get_cbm_defaults_path())()
default_parameters
```
The parameters can be changed here, and will set the value during simulation via
a parameter factory function

```{r}
# replace the slow mixing rate parameter with a new value
default_parameters$slow_mixing_rate[,] <- 0.005

# create a new "cbm_parameters_factory" with the customizations
parameter_factory <- function(){
  return(default_parameters)
}
```

CBM Spinup function - initialize the CBM pools and state (cbm_vars)

```{r}


spinup <- function(
  cbm_factory,
  parameter_factory,
  inventory,
  classifiers,
  merch_volumes
){
  with(cbm_factory$initialize_cbm(
    # note we pass the customized parameter factory here
    cbm_parameters_factory=parameter_factory
  ) %as% cbm, {

      inv_tuple  = cbm_factory$prepare_inventory(inventory)
      csets = inv_tuple[[1]]
      inv = inv_tuple[[2]]


      cbm_vars <- cbm_variables$initialize_simulation_variables(
          csets,
          inv,
          cbm$pool_codes,
          cbm$flux_indicator_codes,
          inv$backend_type
      )

      spinup_vars = cbm_variables$initialize_spinup_variables(
          cbm_vars,
          inv$backend_type,
          spinup_params=NULL,
          include_flux=FALSE
      )
      cbm$spinup(spinup_vars, reporting_func = NULL)
      cbm_vars = cbm$init(cbm_vars)

      return(cbm_vars)
  })
}
```

CBM Step function - step the annual process and disturbance routine by 1
timestep

```{r}
step <- function(cbm_factory, parameter_factory, cbm_vars){
  with(cbm_factory$initialize_cbm(
    # note we pass the customized parameter factory here
    cbm_parameters_factory=parameter_factory
  ) %as% cbm, {

      cbm_vars = cbm$step_start(cbm_vars)
      cbm_vars = cbm$step_disturbance(cbm_vars)
      cbm_vars = cbm$step_annual_process(cbm_vars)
      cbm_vars = cbm$step_end(cbm_vars)
      return (cbm_vars)

  })
}

```


Create a `StandCBMFactory` object for initializing the cbm model, and a
`CBMOutput`object to store the time step output.


```{r}
cbm_factory <-stand_cbm_factory$StandCBMFactory(
        classifiers, merch_volumes
      )
output <- cbm_output$CBMOutput(
  classifier_map=dict(cbm_factory$classifier_value_names),
  disturbance_type_map=dict(cbm_factory$disturbance_types)
)

```

CBM spinup

```{r}
cbm_vars = spinup(cbm_factory, parameter_factory, inventory)
output$append_simulation_result(0, cbm_vars)
```

CBM time-stepping

```{r}

for (i in 1:50){

  # set the disturbance type array with randomly drawn disturbance types
  # note the CBM default disturbance type ids are used

  # this is here to illustrate that it's possible to read and inspect and modify
  # the CBM simulation state between timesteps
  disturbance_types = sample(
      c(0L,1L,2L),
      n_stands,
      replace=TRUE,
      prob=c(0.95, 0.03, 0.02)
  )

  cbm_vars$parameters["disturbance_type"]$assign(disturbance_types)

  step(cbm_factory, parameter_factory, cbm_vars)
  # record the end of timestep result
  output$append_simulation_result(i, cbm_vars)
}


```


extract the results data frames

```{r}
pools_output <- output$pools$to_pandas()
flux_output <- output$flux$to_pandas()
state_output <- output$state$to_pandas()
parameters_output <- output$parameters$to_pandas()
area <- output$area$to_pandas()
```

Plot some output

```{r}

aggregated_age = aggregate(
  state_output$age, by=list(timestep=state_output$timestep), FUN=mean
)
colnames(aggregated_age)= c("timestep", "age")
plot(aggregated_age$timestep, aggregated_age$age)
```